### 2011-07-20 Text/paper.tex
Need to resolve the two-dimensional cap contributions by automodel solutions.

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amssymb,amsmath}
\usepackage{bm}
\usepackage{color}
\usepackage{tabularx}
\begin{document}
\section{Motivation}
The mass transfer is a very important characteristic for microchannels  design. It allows to
properly manufacture a microchannel with necessary properties to ensure that chemical
reactions are performed in the best manner. In
what follows the mass transfer associated with the bubble moving in the microchannel through liquid
medium is studied.

A few questions need to be addressed:
\begin{description}
 \item[I] The efficiency of the mass transfer depending on the capillary number. For small capillary
numbers as $Ca<0.6$ for square microchannels there is a vortex in the reference frame moving with
the bubble, which  influences and improves the corresponding mass transfer. 
Note that the flow
patterns strongly depend on the number of parameters as the slug length, the thickness of the
thickness and the existence of the vortex in front of the bubble. There are some available
correlations in literature \cite{bercic-mass,kreutzer-overview} and one of the particular goal of
this work is to study mass transfer dependence on the number of parameters. Another goal is to
validate and to come up with another correlation describing the mass transfer.

 \item[II] Another important quantity is to study the mixing processes inside the slug. One of the
possible characteristics of the mixing is a residence time distribution. Once the step function
of tracer is introduced in the liquid film, then the concentration of tracer with time is
measured at the outlet. The form of the signal and especially its standard deviation is responsible
for mixing. However, from numerical point of view this study requires the repetition of domain a
few times to see the tracer propagation. Thus, we do not currently address mixing with our
simulations.
\end{description}

To address the questions arised we will utilize the results of numerical simulations of flow
obtained
before \cite{kuzmin-binary2d}. The numerical benchmark is presented in Fig. \ref{fig:benchmark}. In
the same figure, one can see the proposed benchmark for the tracer. The
physical
picture for the proposed numerical benchmark is  related to classical experiments conducted by
\citet{bercic-mass}. In their work authors studied gas-liquid mass transfer coefficient by
injecting methane as the gas bubble and measuring the dissolved concentration of methane in water
along the channel. This situation implies the periodic boundary conditions and steady-state motion
of bubbles.

The numerical  procedure of studying mass transfer is presented in the following steps:
\begin{description}
 \item[Flow field] The hydrodynamics for the bubble motion in the periodic domain is obtained
according to the work \cite{kuzmin-binary2d}. The grid resolution is taken to insure grid
independency of results.  
 \item[Bubble reference frame] Once the hydrodynamics is resolved, the mass transfer simulations
are conducted in the moving
reference frame with the bubble. Then the bubble stands still and the flow is coming around the
bubble. We impose a steady concentration of solutant on the surface of the bubble.
\item[Mass transfer] One needs not only perform a Galilean transformation to the reference
frame moving with the bubble but also to
transfer the bubble to the inlet condition. This can insure that the studied phenomenon is a liquid
slug. 
\end{description}
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_new.eps}\\
\\
\\
\\
\includegraphics[width=\textwidth]{Figures/benchmark_periodic.eps}
\caption{The two-dimensional benchmarks for the hydrodynamics (top) and the mass transfer
coefficient (bottom). \label{fig:benchmark}}
\end{figure}

\section{Mass Transfer Correlations}
\label{sec:correlations}
We examine the following three characteristic correlations:
\begin{description}
\item[I]
In what follows we want to address the following empirical correlation for the volumetric mass
transfer coefficient addressed in the recent
publication by \citet{yue-mass}:
\begin{equation}
k_L a=\frac{2}{d_h} \Bigl(\frac{D
U_{\mathrm{bubble}}}{L_{\mathrm{bubble}}+L_{\mathrm{slug}}}\Bigr)^{0.5}
\Bigl(\frac{L_{\mathrm{bubble}}}{L_{\mathrm{bubble}}+L_{\mathrm{slug}}}\Bigr)^{0.3}.
\end{equation}
This correlation is within $10\%$ accuracy for experimental results for the following range of
parameters: bubble velocity $0.4\,\mathrm{m/s}<U_{\mathrm{bubble}}<2\,\mathrm{m/s}$, bubble length
$1.4<\frac{L_{\mathrm{bubble}}}{d_h}<6.3$, slug length $1<\frac{L_{\mathrm{slug}}}{d_h}<3.2$,
hydrodynamic diameter $d_h = 400
\mathrm{\mu m}$.

\item[II]
For long slug lengths in circular capillaries, where the mass transfer from bubble caps is
predominant, the folowing correlation is derived by
\citet{bercic-mass}:
\begin{equation}
k_L a = \frac{0.111
(U_{\mathrm{gas}}+U_{\mathrm{liq}})^{1.19}}{\bigl((1-\varepsilon_{\mathrm{gas}})(L_{\mathrm{bubble}}
+L_ {\mathrm{slug}} )\bigr)^{0.57} },
\end{equation}
where $\varepsilon_{\mathrm{gas}}=\frac{U_{\mathrm{gas}}}{U_{\mathrm{bubble}}}$ is the gas holdup,
$U_{\mathrm{gas}}$ is the superficial gas velocity, and $U_{\mathrm{liq}}$ is the superficial
liquid velocity. The superficial liquid velocity is calculated with the steady state condition in
the middle of the slug as:
\begin{equation}
U_{\mathrm{liq}}=\frac{\oint U \mathrm{d}A}{A}.
\end{equation}
The superficial liquid velocity can be calculated through the calculated bubble volume and gas
holdup. For the two-dimensional case it is represented as:
\begin{equation}
\begin{aligned}
&\varepsilon_{\mathrm{gas}}=\frac{\text{bubble area}}{15 H^2}\\
&U_{\mathrm{gas}}=\varepsilon_{\mathrm{gas}} U_{\mathrm{bubble}}.
\end{aligned}
\end{equation}

Since experiments done by \citeauthor{bercic-mass} were only for one diffusivity coefficient
$D=2\times 10^{-9}\,\mathrm{m^2/s}$, the effective equation should be rescaled as:
\begin{equation}
k_L a = \sqrt{\frac{D}{2\times 10^{-9}\,\mathrm{m^2/s}}}\frac{0.111
(U_{\mathrm{gas}}+U_{\mathrm{liq}})^{1.19}}{\bigl((1-\varepsilon_{\mathrm{gas}})(L_{\mathrm{bubble}}
+L_ {\mathrm{slug}} )\bigr)^{0.57} },
\end{equation}

\item[III]
\citet{vanbaten-circular} performed numerical simulations for circular capillaries by assuming that
there is a separate contribution of the liquid film and the bubble caps to mass transfer. Their
correlation is as:
\begin{equation}
k_L a = \frac{2}{\sqrt{\pi}}\sqrt{\frac{D U_{\mathrm{bubble}}}{(L_{\mathrm{bubble}}-d_h)}}
\frac{4(L_{\mathrm{bubble}}-d_h)}{d_h(L_{\mathrm{bubble}}+L_{\mathrm{slug}})}+2\frac{\sqrt{2}}{\pi}
 \sqrt{\frac{D U_{\mathrm{bubble}}}{d_h}} \frac{4}{(L_{\mathrm{bubble}}+L_{\mathrm{slug}})}.
\end{equation}
This correlation is valid for short contact times, when the mass transfer is important from bubble
caps as well as from the film. Short contact times are defined through Fourier number as:
\begin{equation}
Fo=\frac{D t_{\mathrm{film}}}{\delta^2},
\end{equation}
where $t_{\mathrm{film}}=\frac{L_{\mathrm{film}}}{U_{\mathrm{bubble}}}$ is the time exposure of the
bubble while it propagates through distance $L_{\mathrm{film}}$. The Fourier number is assumed to
be small if $Fo<0.1$. If $Fo>1$ then the liquid film is saturated and the mass transfer occurs from
the bubble caps.

\end{description}

Overall work goal is to compare numerical simulations with given correlations and to establish a
new one.

\section{Mass transfer}
The mass transfer is characterized by non-dimensional numbers as Schmidt number (viscous diffusion
rate over mass diffusion rate), the Peclet number (convection over diffusion) and the Sherwood
number (convection mass transfer over diffusion mass transfer):
\begin{equation}
\begin{aligned}
&Pe=\frac{U L}{D}\\
&Sh=\frac{K L}{D}\\
&Sc=\frac{\nu}{D}\\
\end{aligned}
\end{equation}

As far as the lattice Boltzmann method operates only in the non-dimensional space, one needs to
match the non-dimensional lattice Boltzmann parameters with non-dimensional physical parameters.
Therefore, one needs to form non-dimensional numbers from the physical parameters. One of the
particular experiment examples is the
work by \citet{bercic-mass} who used the methane dissolution to characterize the mass transfer for
bubble train flow.

In what follows we will perform the matching of obtained hydrodynamics results
\cite{kuzmin-binary2d} with physical ones from experiment of \citet{bercic-mass}. The diffusion
coefficient of the methane under standard conditions \cite{methane-properties} is
$1.84\times 10^{-5} \mathrm{cm}^2/\mathrm{s}$. The diameters of capillaries in the experiment were
as $1.5$, $2.5$ and
$3.1\,\mathrm{mm}$. For simplicity reasons we will choose the diameter of capillary as $1.5\,
\mathrm{mm}$. 
The non-dimensional parameters obtained from the hydrodynamics lattice Boltzmann simulations are
as:
\begin{equation}
\begin{aligned}
&Ca=\frac{U \mu_{\mathrm{liq}}}{\gamma}\\
&Re=\frac{\rho_{\mathrm{liq}} U L}{\mu_{\mathrm{liq}}}.
\end{aligned}
\end{equation}
Given some physical characteristics as the liquid density and the channel length,
one can obtain the physical velocity obtained from non-dimensional numbers:
\begin{equation}
\begin{aligned}
Ca\cdot Re= \frac{\rho_{\mathrm{liq}} U^2 L}{\gamma}\\
U=\sqrt{\frac{Ca\,Re\,\gamma}{\rho_{\mathrm{liq}}L}}
\end{aligned}
\end{equation}

Assuming that the surface tension of water is $0.0728\,\mathrm{N/m}$ (methane exhibits the same
trend \cite{schmidt-methane}), the density is       
$1000\,\mathrm{kg/m^3}$, the hydraulic diameter of the channel to be $1.5\,\mathrm{mm}$ one can
calculate the physical velocity of the bubble.
Table \ref{table:twod:simulations} represents lattice Boltzmann simulations results and their
physical equivalents for the grid $202\times 3002$. One can see
that velocity is quite reasonable
as \citet{bercic-mass} declared that experiments were conducted for velocities in the ranges of
$0.01$ to $0.4\,\mathrm{m}/\mathrm{s}$. 
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$Ca$    &$Re$     &$U_{LB}$ &$\delta$&$\varepsilon_{\mathrm{gas}}$
&$U_{\mathrm{phys}}, \mathrm{m/s}$\\
\hline
$0.026$ &$0.449$  &$0.0014$ &$0.040$ &$0.306$ &$0.023$   \\ 
$0.047$ &$0.820$  &$0.0027$ &$0.058$ &$0.293$ &$0.043$   \\ 
$0.080$ &$1.378$  &$0.0045$ &$0.085$ &$0.280$ &$0.073$ \\
$0.065$ &$1.126$  &$0.0037$ &$0.076$ &$0.266$ &$0.059$      \\
$0.222$ &$3.807$  &$0.0125$ &$0.122$ &$0.253$ &$0.202$  \\
$0.479$ &$8.222$  &$0.0271$ &$0.151$ &$0.249$ &$0.437$  \\
$0.736$ &$12.617$ &$0.0416$ &$0.164$ &$0.236$ &$0.671$  \\ 
$0.989$ &$16.960$ &$0.0559$ &$0.172$ &$0.230$ &$0.902$  \\
\hline
\end{tabularx}
\caption{Two-dimensional simulations for different
capillary numbers. \label{table:twod:simulations}}
\end{table}

\section{Lattice Boltzmann implementation}
For the lattice Boltzmann implementation we took $\omega=1.99$ which results in the following
lattice Boltzmann diffusivity parameter:
\begin{equation}
\begin{aligned}
&D=\frac{1}{3}\Bigl(\frac{1}{\omega}-\frac{1}{2}\Bigr)\\
&D=0.0008375
\end{aligned}
\end{equation}
In terms of the Schmidt number (taking the lattice Boltzmann liquid viscosity in former simulations
as $\frac{2}{3}$) that
can result in:
\begin{equation}
\begin{aligned}
&Sc=\frac{\nu_{\mathrm{liq}}}{D}\\
&Sc=796\\
\end{aligned}
\end{equation}
Taking the water dynamic viscosity as $8.9 \times 10^{-4}\,\mathrm{Pa\cdot s}$ or kinematic
viscosity as $8.9 \times
10^{-7} \,\mathrm{m^2/s}$ and the Schmidt number as $796$ results in the following physical
diffusion coefficient:
\begin{equation}
D=1.118\times 10^{-5}\,\mathrm{cm^2/s},
\end{equation}
which is less than the diffusivity of the methane $1.84\times 10^{-5}\,\mathrm{cm^2/s}$ but is
physical. After calculation of physical parameters one can calculate correlations indicated in
Section \ref{sec:correlations} for mass transfer. Table \ref{table:lengths} presents extraction
from the lattice Boltzmann simulation necessary to calculate mass transfer calculations.
For the two-dimensional case parameter  $a$ is defined as:
\begin{equation}
a = \frac{\text{bubble perimeter}}{\text{cell area}}.
\end{equation}
For the height of the channel to be $H=1.5 \times 10^{-3}\,\mathrm{m}$, parameters $a$ are
presented in Table \ref{table:lengths}. 
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$Ca$&Bubble length& Slug Length&
$U_{\mathrm{gas}},\,\mathrm{m/s}$&$U_{\mathrm{liq}},\,\mathrm{m/s}$&$a,\,\mathrm{m^{-1}}$\\
\hline
$0.026$&$5.225$&$9.775$ &$0.0070$ &$0.0209$&$510.77$\\
$0.047$&$5.22$ &$9.78$  &$0.0126$ &$0.0379$&$508.62$\\
$0.080$&$5.215$&$9.785$ &$0.0204$ &$0.0615$&$506.53$\\
$0.065$&$4.965$&$10.035$&$0.0157$ &$0.0504$&$482.84$\\
$0.222$&$5.25$ &$9.75$  &$0.0511$ &$0.1560$&$505.06$\\
$0.479$&$5.565$&$9.435$ &$0.1089$ &$0.3155$&$530.37$\\
$0.736$&$5.51$ &$9.49$  &$0.1584$ &$0.4677$&$525.38$\\
$0.989$&$5.505$&$9.495$ &$0.2074$ &$0.6190$&$526.80$\\
\hline
\end{tabularx}
\caption{The nondimensional (scaled to the height of the channel) lengths for the same
lattice Boltzmann simulations as in Table \ref{table:twod:simulations}. As well superficial
gas and liquid velocities are presented. \label{table:lengths}} 
\end{table}

The literature correlations of mass transfer dependence on the bubble velocities are presented in
Fig. \ref{fig:mass:transfer:theoretical}. Note, that the correlations presented in
Fig. \ref{fig:mass:transfer:theoretical} are used primarily for circular capillaries
(\citeauthor{vanbaten-circular,bercic-mass}) and for square capillaries (\citeauthor{yue-mass}). We
adapt these results for flow between parallel plates by using hydraulic
diameter \cite{bercic-mass} which can be calculated as:
\begin{equation}
d_h = \lim_{W\to\infty}\frac{4A}{P}=\lim_{W\to\infty}\frac{4 W\,H}{2(W+H)}=2 H.
\end{equation}
For comparison reasons we as well present the correlation when the hydraulic diameter is taken as
$d_h=H$, see Fig. \ref{fig:mass:transfer:theoretical}.

\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/theoretical_correlations.eps}\\
\includegraphics[width=\textwidth]{Figures/theoretical_correlations_without_hydraulic.eps}\\
\caption{The literature correlations by various authors with hydraulic diameter taken as $2 H$
(top) and $H$ (bottom). Correlations deviate a lot among each
other.\label{fig:mass:transfer:theoretical}}
\end{figure}

The volumetric mass transfer coefficient is calculated according to the work of
\citet{vanbaten-circular}. The concentration flux is calculated as the difference between overall
average concentration taken in the whole domain ($C_{overall}=\int_{V} C \mathrm{d}V /V$)
at time
$t_1$ and time $t_2$ divided on the time difference $t_2-t_1$ {\color{red} (In this particular
moment I don't know whether to take the volume of the whole domain or
just the liquid phase. I will go for whole domain.)}. Then the volumetric mass transfer
coefficient is calculated as \cite{vanbaten-circular}:
\begin{equation}
\label{main:simulation:equation}
k_L a=\frac{\text{Flux}}{C_s-C_{outlet}} \frac{\text{bubble surface area}}{\text{unit cell volume}},
\end{equation}
where $C_{outlet}=\int{C U_{outlet}\mathrm{d}A}/\int{U_{outlet}\mathrm{d}A}$. 
{\color{red} Do not understand why we need this ratio bubble surface area/unit cell volume. In this
case the dimensional units do not coincide. In the calculations I will avoid it until further
clearance.} The overall concentration in the lattice Boltzmann system is prescribed as:
\begin{equation}
C_{overall}=\frac{\sum{c_i}}{N_x \, (N_y-2)}.
\end{equation}

To properly do
non-dimensioanalization one needs to insure the time conversion factor when the flux is calculated
inside the lattice Botlzmann system. The time conversion can be easily obtained:
\begin{equation}
\begin{aligned}
&U_{LB}=U_{phys}\frac{\Delta t}{\Delta x}\\
&\Delta t=\frac{U_{LB}}{U_{phys}}\Delta x\\
&\Delta t=\frac{1.5\,\mathrm{mm}}{200} \frac{0.0559}{0.902\,\mathrm{m/s}}=4.64\times
10^{-7}\,\mathrm{s} 
\end{aligned}
\end{equation}

The bubble surface length is the perimeter of the bubble in two-dimensional space and it is
calculated with the extraction of the bubble contour and representing it with the polygon. After
that the calculated perimeter is the effective summation of line lengths and can be represented as:
\begin{equation}
P=\sum_i{\sqrt{(x2_i-x1_i)^2+(y2_i-y1_i)^2}},
\end{equation}
where the index $i$ represents the polyline consisting of two points $(x1,y1)$ and $(x2,y2)$.
%The corresponding volumetric mass transfer coefficients depending on the velocity are indicated in
%Table \ref{table:volumetric:coefficients}.

\section{Simulation results}
\subsection{Steady-state}
Using Eq.\ref{main:simulation:equation} we calculated the mass transfer coefficient. One can see
the plots for the mass transfer coefficient versus time in Fig. \ref{fig:steady:state}. We state as
a steady-state condition the average of concentration values between $500000$ to $1000000$ time
steps.{\color{red} We need to put more time steps. As well it is probably connected with Fourier
number.} The average calculated volumetric mass transfer coefficients values are presented in Table
\ref{table:volumetric:coefficients}. For each of the measurements we as well provided the Fourier
number (for the flow between plates is as follows):
\begin{equation}
Fo=\frac{D t_{\mathrm{film}}}{\delta^2}=\frac{D
(L_{\mathrm{bubble}}-H)}{U_{\mathrm{bubble}}\delta^2}.
\end{equation}
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/steady_state.eps}
\caption{The mass transfer coefficient vs. time iterations. One can see that the pseudo steady
state can be reached after $500000$. The mass transfer coefficient is taken as the average of
concentration values between $500000$ to $1000000$ time steps.\label{fig:steady:state}}
\end{figure}
\begin{table}[htb!]
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
$Ca$&$Fo$&$k_L a,\,\mathrm{s}^{-1}$\\
\hline
$0.026$&$0.08493$&$0.2737$\\
$0.047$&$0.02160$&$0.3664$\\
$0.080$&$0.00588$&$0.4959$\\
$0.065$&$0.00862$&$0.4734$\\
$0.222$&$0.00103$&$0.8505$\\
$0.479$&$0.00034$&$1.1541$\\
$0.736$&$0.00018$&$1.1801$\\
$0.989$&$0.00012$&$1.2215$\\
\hline
\end{tabularx}
\caption{Simulation results for different capillary numbers. \label{table:volumetric:coefficients}}
\end{table}

Comparison between correlations and simulations one can find in Fig.
\ref{fig:comparison:correlations}.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/comparison_correlations.eps}
\caption{One can see the comparison between correlations and simulations. {\color{red} Last two
points can be eliminated due to improper imposement of the concave structure of the bubble. The
large deviation of the results can be probably explained by the flow between plates in comparison
with the flow in circular and square microchannels.} \label{fig:comparison:correlations}}
\end{figure}

\appendix

\section{Boundary conditions}
In what follows we will examine different LBM implementations:
\begin{description}
 \item[BB conditions] The walls can be treated via the simple bounce back rule:
\begin{equation}
f_{\bar{i}B}=f_{iF},\text{ for } \forall i,
\end{equation}
where $B$ is the boundary node with coordinates $\bm{r_F}+\bm{c_i}$, $F$ is the fluid node. This
will fulfill the macroscopic conditions as:
\begin{equation}
J_{fluid}= \sum_{c_i\cdot n>0}{f_i c_i}=J_{wall},
\end{equation}
which is in turn represents the following boundary condition:
\begin{equation}
\partial_n J_n=\partial_n C u_n = u_n \partial_n C=0
\end{equation}

The walls can be treated through simmetric boundary conditions as well, which are really close to
bounce-back formulation. In this case one can insure that the concentration values at the wall and
at the fluid equal to each other.

For the constant concentration at the bubble surface one can use the pressure anti
bounce-back conditions
\cite{ginzburg-boundary-main}:
\begin{equation}
f_{iB}=-f^{*}_{iF}+2 e_i(W),
\end{equation}
where the subscript $*$ stands for after-collision population, $W$ stands for the wall
concentration to be imposed.
\item[Inamuro conditions] Inamuro boundary conditions impose strict conservation of the population
by utilizing diffuse interface kinetic boundary approach \cite{inamuro-scalar-boundary}. The
unknown populations $f_i=w_i \rho^{\prime}\text{ for } \bm{c_i}\cdot \bm{n}>0$ are prescribed
through the equilibrium conditions, where $\rho^{\prime}$ is an unknown parameter.Then the
prescribed wall condition is as: 
\begin{equation}
\rho^{\prime}=\frac{\rho_{w}-\sum_{i,\bm{c_i}\cdot\bm{n}\leq0}
f_i}{\sum_{i,\bm{c_i}\cdot\bm{n}>0}w_i}
\end{equation}
For the zero flux on the wall one can use another equation:
\begin{equation}
\rho^{\prime}=-\frac{i,\bm{c_i}\cdot\bm{n}\leq0}{\sum_{i,\bm{c_i}\cdot \bm{n}}{w_i \bm{c_i}\cdot
\bm{n}}}.
\end{equation}
\end{description}

\section{TRT $D2Q9$ model. Equilibrium functions}
One of the advantages of $D2Q9$ TRT model in comparison with $D2Q5$ usually used for simulations of
the advection-diffusion model is that $D2Q9$ is able to reduce numerical diffusion from diagonal
and non-diagonal diffusion tensor components \cite{kuzmin-stability-optimal}.

In
comparison with the BGK collision operator the TRT collision operator \cite{ginzburg-boundary-main}
decomposes the populations and the equilibrium
distribution into the symmetric and antisymmetric parts:
\begin{equation}
\label{trtdecomp}
f^{\pm}_i=\frac{f_i\pm f_{\bar{i}}}{2}\;,\; 
{eq_i}^{\pm}=\frac{eq_i\pm eq_{\bar{i}}}{2}\;,
\end{equation}
where $\bar{i}$ is the opposite direction to the $i$-th direction.
The collision is performed with two independent relaxation rates for 
symmetric and antisymmetric modes:
\begin{equation}
\label{trt}
\begin{aligned}
&f_i^{*}(\bm{x},t)=f_i(\bm{x},t)-\omega_{+} (f_i^{+} - eq_i^+)-\omega_{-}
(f_i^{-} -
eq_i^-)\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t).
\end{aligned}
\end{equation}
Note that the BGK collision operator is the particular subclass of the TRT relaxation operator with
$\omega_{+}=\omega_{-}$. In comparison with the BGK collision operator,
the TRT collision operator has the additional degree of freedom. Thus, the TRT operator then
introduces
so-called magic (free) parameter
$\Lambda=\Bigl(\frac{1}{\omega_{+}}-\frac{1}{2}\Bigr)\Bigl(\frac{1}{\omega_{-}}-\frac{1}{2}\Bigr)$. 
This free parameter controls the effective location of the bounce-back
walls \cite{ginzburg-multireflection}, second-order accuracy of
boundary \cite{ginzburg-boundary-main} and interface schemes \cite{ginzburg-discontinious}, 
spatial accuracy \cite{ginzburg-recurrence,servan-trt-stability},
consistency \cite{ginzburg-brinkman} and, to some extent,
stability \cite{kuzmin-stability-optimal,kuzmin-d1q3,servan-trt-stability}.
In particular, $\Lambda=\frac{1}{4}$ achieves the optimal stability for the
linear advection-diffusion isotropic equation \cite{kuzmin-stability-optimal}.

The equilibrium functions for $D2Q9$ TRT model are represented as \cite{kuzmin-stability-optimal}:
\begin{equation}
\begin{aligned}
&eq_i^{+}=eq_i^{(m)}+g^{(u)} eq_i^{(u)}\\
&eq_i^{(m)}=t_i^{(m)} c_e+ eq_i^{(a)}\\
&eq_i^{(u)}=t_i^{(u)} \frac{u_x^2+u_y^2}{2}+\frac{u_x^2-u_y^2}{4} p_i^{(xx)}+g_{xy}^{(u)}\frac{u_x
u_y}{4} p_i^{xy}\\
&eq_i^{(a)}=\frac{D_{xx}-D_{yy}}{4} p_i^{xx}+\frac{D_{xy}}{4} p_i^{(xy)},
\end{aligned}
\end{equation}
where $p_i^{(xx)}=c_{ix}^2-c_{iy}^2$ and $p_i^{(xy)}=c_{ix} c_{iy}$.

The following equilibrium functions solve the anisotropic advection-diffusion equation with the
following diffusion tensor:
\begin{equation}
D=
\begin{pmatrix}
D_{xx} + (g^{(u)}-1) u_x^2 & D_{xy}+(g_{xy}^{(u)}-1)u_x u_y\\
D_{xy} + (g_{xy}^{(u)}-1) u_x u_y& D_{yy}+(g^{(u)}-1) u_y^2 
\end{pmatrix}
\end{equation}
To consider the situation when the resolution is not needed to be changed in $x$ and $y$
directions, one can obtain that $D_{xx}=D_{yy}$. The cross-diagonal components of the diffusion
tensor are taken as zeros, i.e. $D_{xy}=0$. The numerical diffusion can be avoided by proper choice
of the equilibrium functions, i.e. $g_{xy}^{(u)}=g^{(u)}=1$.

As far as the simulated diffusion tensor is close to $0$ the stability limits restrict the velocity
amplitude to be simulated \cite{kuzmin-stability-optimal}. However, particular choice of weights
can drastically help. Fig. \ref{stability:d2q9} shows the stability limits for D2Q9 with removed
numerical diffusion and with it for two particular choices of weights.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/d2q9_stability.eps}
\caption{The non-negativity areas and stable sub-domains of the OTRT schemes ($\Lambda=\frac{1}{4}$)
are plotted for the $D2Q9$ (standard weights, left) and the $D2Q9$ (uniform weights, right) when the
numerical diffusion is removed, partially or completely. Two non-negativity areas where all the
equilibrium populations are positive have curvilinear boundaries: they are smaller when
$g^{(u)}_{xy}=0$ (yellow) and than when $g^{(u)}_{xy}=1$ (green+yellow). The non-negativity
condition is required by BGK collision operator. The stable sub-domains for $g^{(u)} = 1,
g^{(u)}_{xy} = 0$ are small (red+green+yellow) triangles. When the numerical diffusion is removed
the stable area is bounded by magenta and the (blue) domain above. The proper choice of weights can
allow to obtain  total triangle $0 \leq u_x^2+u_y^2 \leq 1−c_e$ (the
uniform weights only).
\label{stability:d2q9}}
\end{figure}

From here on we will use the standard weights as:
\begin{equation}
t_i^{(u)}=t_i^{(m)}=t_i^{(a)}=\Bigl\{0,\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{12},
\frac {1}{12},\frac{1}{12},\frac{1}{12}\Bigr\}
\end{equation}

\bibliographystyle{unsrtnat}
\bibliography{paper}

\end{document}